Traffic Aware Resource Allocation Schemes for Multi-Cell \acs{MIMO}-\acs{OFDM} Systems

### 2015-03-02 Report/convergence_centralized.tex
updated

When the objective function is zero, the uniqueness of the transmit precoders using \eqref{uniqueness-1} and \eqref{uniqueness-2} are not valid due to the inactive constraints \eqref{eqn-6.2} and \eqref{eqn-mse-1.2}. To obtain a unique set of transmit precoders when the objective is zero even for a single \ac{BS}, we can regularize the objective with the total transmit power expression without affecting the optimal solution as 
\begin{equation*}
\| \tilde{\mbf{v}} \|_q + \varphi \sum_{k \in \mc{U}} \sum_{n = 1}^{N} \sum_{l=1}^{L} \mathrm{tr} \left ( \mvec{m}{l,k,n} \mbf{m}^\herm_{l,k,n} \right ),
\end{equation*}
for some arbitrarily small value of the scaling factor \me{\varphi \approx 0}.

@@

\begin{comment}

To prove the convergence of the centralized algorithm in \eqref{eqn-6} and \eqref{eqn-mse-1}, following conditions are necessary to prove the sequence convergence using Bolzano-Weierstrass theorem \cite{rudin1964principles}.
\begin{itemize}
	\item At each \ac{SCA} update, the objective value decreases monotonically 
	\item At each \ac{AO} update, the objective value should decrease monotonically
	\item Boundedness of the objective value sequence and the uniqueness of the limiting point.
\end{itemize}
In addition, we show that the converged point is in fact the stationary point of the original nonconvex problem in \eqref{eqn-6} and \eqref{eqn-mse-1}.

>>>>

When the objective function is zero, the uniqueness of the transmit precoders using \eqref{uniqueness-1} and \eqref{uniqueness-2} are not valid due to the inactive constraints \eqref{eqn-6.2} and \eqref{eqn-mse-1.2}. To obtain a unique set of transmit precoders when the objective is zero even for a single \ac{BS}, we can follow the regularization approach discussed in Appendix \ref{a-3} without affecting the optimal solution.
%we can regularize the objective with the total transmit power expression without affecting the optimal solution as 
%\begin{equation*}
%\| \tilde{\mbf{v}} \|_q + \varphi \sum_{k \in \mc{U}} \sum_{n = 1}^{N} \sum_{l=1}^{L} \mathrm{tr} \left ( \mvec{m}{l,k,n} \mbf{m}^\herm_{l,k,n} \right ),
%\end{equation*}
%for some arbitrarily small value of the scaling factor \me{\varphi \approx 0}.

@@

\begin{comment}
To prove the convergence of the centralized algorithm in \eqref{eqn-6} and \eqref{eqn-mse-1}, following conditions are necessary to prove the sequence convergence using Bolzano-Weierstrass theorem \cite{rudin1964principles}.
\begin{itemize}
	\item At each \ac{SCA} update, the objective value decreases monotonically 
	\item At each \ac{AO} update, the objective value should decrease monotonically
	\item Boundedness of the objective value sequence and the uniqueness of the limiting point.
\end{itemize}
In addition, we show that the converged point is in fact the stationary point of the original nonconvex problem in \eqref{eqn-6} and \eqref{eqn-mse-1}.

### 2015-03-02 Report/introduction.tex
updated

The queue minimizing precoder designs are closely related to the \ac{WSRM} problem with additional rate constraints determined by the number of backlogged packets for each user in the system. The topics on \ac{MIMO} \ac{IBC} precoder design have been studied extensively with different performance criteria in the literature. Due to the nonconvex nature of the \ac{MIMO} \ac{IBC} precoder design problems, the \ac{SCA} method has become a powerful tool to deal with these problems \cite{bertsekas1999nonlinear}. For example, in \cite{sin_algorithm}, the nonconvex part of the objective has been linearized around an operating point in order to solve the \ac{WSRM} problem in an iterative manner. Similar approach of solving the \ac{WSRM} problem by using arithmetic-geometric inequality has been proposed in \cite{tran2012fast}.

@@

\subsubsection*{Main Contributions}
\review{In this paper, we design the precoders jointly across space-frequency resources by minimizing the total number of backlogged packets waiting at the \acp{BS} queues. Since the transmissions are guided by the backlogged packets, the proposed formulation limit the allocations beyond the number of backlogged packets without additional rate constraint. We adopt \ac{SCA} to solve by a sequence of convex subproblems using first order approximations due to the nonconvexity imposed by the \ac{DC} constraints.} Initially, we propose centralized \ac{JSFRA} algorithms, which employs \ac{SCA} for the nonconvex \ac{DC} constraint. First method is by using the direct formulation and the second one is by using the \ac{MSE} equivalence with the rate expression to solve for an optimal precoders. Then we propose distributed precoder designs based on the primal and the \ac{ADMM} methods. Finally, we propose a iterative practical algorithm to decouple the precoder design across the coordinating \acp{BS} with limited information exchange by solving the \ac{KKT} conditions for the \ac{MSE} reformulation solution. \review{It is worth noting that the joint space-frequency channel matrix can be formed by stacking the channel of each sub-channel in a block-diagonal form for all users}.

>>>>

The queue minimizing precoder designs are closely related to the \ac{WSRM} problem with additional rate constraints determined by the number of backlogged packets for each user in the system. The topics on \ac{MIMO} \ac{IBC} precoder design have been studied extensively with different performance criteria in the literature. Due to the nonconvex nature of the \ac{MIMO} \ac{IBC} precoder design problems, the \ac{SCA} approach has become a powerful tool to deal with these problems. For example, in \cite{sin_algorithm}, the nonconvex part of the objective has been linearized around an operating point in order to solve the \ac{WSRM} problem in an iterative manner. Similar approach of solving the \ac{WSRM} problem by using arithmetic-geometric inequality has been proposed in \cite{tran2012fast}.

@@

\subsubsection*{Main Contributions}
\review{In this paper, we design the precoders jointly across space-frequency resources by minimizing the total number of backlogged packets waiting at the \acp{BS} queues. Since the transmissions are guided by the backlogged packets, the proposed formulation limit the allocations beyond the number of backlogged packets without additional rate constraint. We adopt \ac{SCA} to solve by a sequence of convex subproblems using first order approximations due to the nonconvexity imposed by the \ac{DC} constraints.} Initially, we propose centralized \ac{JSFRA} algorithms, which employs \ac{SCA} for the nonconvex \ac{DC} constraint. First method is by using the direct formulation and the second one is by using the \ac{MSE} equivalence with the rate expression to solve for an optimal precoders. Then we propose distributed precoder designs based on the primal and the \ac{ADMM} methods. Finally, an iterative practical algorithm is proposed to decouple the precoder design across the coordinating \acp{BS} with limited information exchange by solving the \ac{KKT} conditions for the \ac{MSE} reformulation. \review{Note that the joint space-frequency channel matrix can be formed by stacking the channel of each sub-channel in a block-diagonal form for all users}.

### 2015-03-02 Report/wsrm_formulation.tex
updated

We now present an iterative algorithm to solve problem \eqref{eqn-3} by using alternating optimization technique in conjunction with \acf{SCA} \cite{marks1978technical}. The problem is to determine the transmit precoders \me{\mvec{m}{l,k,n}} and the receive beamformers \me{\mvec{w}{l,k,n}} to minimize the total number of backlogged packets in the system. Since the \ac{SINR} expression in \eqref{eq:SINR} cannot be handled directly to formulate the problem, we relax the equality constraint of \eqref{eq:SINR} using two inequality constraints by introducing additional variables to solve \eqref{eqn-3} as
\iftoggle{single_column}{
\begin{IEEEeqnarray}{CCl}\label{eqn-6}
\underset{\substack{\gamma_{l,k,n},\mvec{m}{l,k,n},\\ \beta_{l,k,n},\mvec{w}{l,k,n}}}{\text{minimize}} & \quad & \|  \tilde{\mbf{v}}  \|_q \label{eqn-obj} \IEEEyessubnumber \\
\text{subject to}& \quad &\gamma_{l,k,n} \leq \frac{ | \mvec{w}{l,k,n}^\herm \mvec{H}{l,k,n} \mvec{m}{l,k,n}  |^2}{\beta_{l,k,n}} \IEEEyessubnumber \label{eqn-6.2} \\
  & \quad & \beta_{l,k,n} \geq  \enoise + \hspace{-0.75em} \sum_{(j,i) \neq (l,k)} \hspace{-0.75em} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2 \IEEEyessubnumber \label{eqn-6.3} \\
  & \quad & \sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \sum_{l=1}^L \trace \, (\mvec{m}{l,k,n} \mvec{m}{l,k,n}^\herm) \leq P_{{\max}}, \fall b. \IEEEyessubnumber \label{eqn-6.4}
\end{IEEEeqnarray}
}{\allowdisplaybreaks
\begin{IEEEeqnarray}{CCl} \label{eqn-6}
	\underset{\substack{\gamma_{l,k,n},\mvec{m}{l,k,n},\\\beta_{l,k,n},\mvec{w}{l,k,n}}}{\text{minimize}} &\quad& \|  \tilde{\mbf{v}}  \|_q \label{eqn-obj} \IEEEyessubnumber \vspace{-0.25cm}\\
	\text{subject to} &\quad& \gamma_{l,k,n} \leq \frac{ | \mvec{w}{l,k,n}^\herm \mvec{H}{l,k,n} \mvec{m}{l,k,n}|^2}{\beta_{l,k,n}} \IEEEyessubnumber \eqspace \label{eqn-6.2} \\
	&\quad& \beta_{l,k,n} \geq \enoise + \sum_{\mathclap{(j,i) \neq (l,k)}} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2 \IEEEyessubnumber \eqspace \label{eqn-6.3} \\
	&\quad& \sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \sum_{l=1}^L \trace \, (\mvec{m}{l,k,n} \mvec{m}{l,k,n}^\herm) \leq P_{{\max}}, \fall b. \IEEEyessubnumber \eqspace \label{eqn-6.4}
\end{IEEEeqnarray}
}
\review{The \ac{SINR} expression in \eqref{eq:SINR} is relaxed by the inequalities \eqref{eqn-6.2} and \eqref{eqn-6.3} in the above formulation. Note that \eqref{eqn-6.2} is an under estimator for \ac{SINR} \me{\gamma_{l,k,n}}, and \eqref{eqn-6.3} provides an upper bound for the total interference seen by user \me{k \in \mathcal{U}_b}, denoted by the variable \me{\beta_{l,k,n}}}. \review{The constraints are tight when each \ac{BS} objective is non zero at the optimal point, as discussed in Appendix \ref{a-3}}. \review{Since the \ac{JSFRA} formulation can be modeled as a \ac{WSRM} problem, which is known to be NP-hard \cite{np_hard}, it also belongs to the class of NP-hard problems.}

>>>>

We present an iterative algorithm to solve \eqref{eqn-3} by using alternating optimization technique in conjunction with \acf{SCA} \cite{marks1978technical}. The problem is to determine the transmit precoders \me{\mvec{m}{l,k,n}} and the receive beamformers \me{\mvec{w}{l,k,n}} to minimize the total number of backlogged packets in the system. The \ac{SINR} expression in \eqref{eq:SINR} cannot be used to formulate the problem directly due to the equality constraint of \eqref{eq:SINR}. Using additional variables, we relax the expression \eqref{eq:SINR} by inequality constraints to solve \eqref{eqn-3} as
\iftoggle{single_column}{
\begin{IEEEeqnarray}{CCl}\label{eqn-6}
\underset{\substack{\gamma_{l,k,n},\mvec{m}{l,k,n},\\ \beta_{l,k,n},\mvec{w}{l,k,n}}}{\text{minimize}} & \quad & \|  \tilde{\mbf{v}}  \|_q \label{eqn-obj} \IEEEyessubnumber \\
\text{subject to}& \quad &\gamma_{l,k,n} \leq \frac{ | \mvec{w}{l,k,n}^\herm \mvec{H}{l,k,n} \mvec{m}{l,k,n}  |^2}{\beta_{l,k,n}} \IEEEyessubnumber \label{eqn-6.2} \\
  & \quad & \beta_{l,k,n} \geq  \enoise + \hspace{-0.75em} \sum_{(j,i) \neq (l,k)} \hspace{-0.75em} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2 \IEEEyessubnumber \label{eqn-6.3} \\
  & \quad & \sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \sum_{l=1}^L \trace \, (\mvec{m}{l,k,n} \mvec{m}{l,k,n}^\herm) \leq P_{{\max}}, \fall b. \IEEEyessubnumber \label{eqn-6.4}
\end{IEEEeqnarray}
}{\allowdisplaybreaks
\begin{IEEEeqnarray}{CCl} \label{eqn-6}
	\underset{\substack{\gamma_{l,k,n},\mvec{m}{l,k,n},\\\beta_{l,k,n},\mvec{w}{l,k,n}}}{\text{minimize}} &\quad& \|  \tilde{\mbf{v}}  \|_q \label{eqn-obj} \IEEEyessubnumber \vspace{-0.25cm}\\
	\text{subject to} &\quad& \gamma_{l,k,n} \leq \frac{ | \mvec{w}{l,k,n}^\herm \mvec{H}{l,k,n} \mvec{m}{l,k,n}|^2}{\beta_{l,k,n}} \IEEEyessubnumber \eqspace \label{eqn-6.2} \\
	&\quad& \beta_{l,k,n} \geq \enoise + \sum_{\mathclap{(j,i) \neq (l,k)}} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2 \IEEEyessubnumber \eqspace \label{eqn-6.3} \\
	&\quad& \sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \sum_{l=1}^L \trace \, (\mvec{m}{l,k,n} \mvec{m}{l,k,n}^\herm) \leq P_{{\max}}, \fall b. \IEEEyessubnumber \eqspace \label{eqn-6.4}
\end{IEEEeqnarray}
}
\review{The \ac{SINR} expression in \eqref{eq:SINR} is relaxed by the inequalities \eqref{eqn-6.2} and \eqref{eqn-6.3} in the above formulation. Note that \eqref{eqn-6.2} is an under estimator for \ac{SINR} \me{\gamma_{l,k,n}}, and \eqref{eqn-6.3} provides an upper bound for the total interference seen by user \me{k \in \mathcal{U}_b}, denoted by the variable \me{\beta_{l,k,n}}}. \review{The constraints are tight when each \ac{BS} objective is non zero at the optimal point, as discussed in Appendix \ref{a-3}}. \review{Since the \ac{JSFRA} formulation can be modeled as a \ac{WSRM} problem, which is known to be NP-hard \cite{np_hard}, it also belongs to the class of NP-hard problems.}
